---
title: "CEDARO report: Behavioural and clinical predictors for Loiasis"

header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output:
  
  pdf_document: default
  html_document: default
  number_sections: true

---

# Background

The “CERMEL Data Repository” (CEDARO) aims to archive data of all research projects performed at
CERMEL. A detailed description of the repository is available from the CEDARO team (cedaro@cermel.org).
Before archiving, a codebook and an anonmyized dataset are generated. In addition, a basic statistical analysis
is performed to ensure that the data in the publication corresponds to the submitted dataset. Typically, only
data to ensure that the numbers of the basic characteristics of the study population and of the outcomes are correct. Rarely would effect sizes, regression models etc. be verified.
A report comparing the original published results with that of the CEDARO analysis is generated and sent
to the investigator who submitted the data set. The CEDARO team will follow-up on any descripancies to
make sure that the analysis and data are valid.

# Publication

**Reference**

Mischlinger J, Veletzky L, Tazemda-Kuitsouc GB, Pitzinger P, Matsegui PB,
Gmeiner M, Lagler H, Gebru T, Held J, Mordmüller B, Ramharter M. Behavioural and clinical predictors for Loiasis. J Glob Health. 2018 Jun;8(1):010413. 

DOI: 10.7189/jogh.08.010413

PubMed PMID : 29497506 

PubMed-Central PMCID: PMC5827628

**Abstract**

"Loiasis is a vector-borne disease in Central and West Africa. While
there is still uncertainty to what extent loiasis is responsible for population
morbidity, individuals having both loiasis and onchocerciasis have a high risk
of fatal encephalopathy when treatment (ie, ivermectin) for onchocerciasis is
given. Therefore it is current policy that communities of high loiasis-burden
are excluded from mass drug administration programmes of ivermectin. To
address this treatment gap we present diagnostic scores, based on clinical and
behavioural predictors that may help to rapidly identify sub-groups with loiasis within high-burden communities."

```{r echo = FALSE, warning = FALSE, message = FALSE}


library(tidyverse)
library(readstata13)
library(knitr)

```
# Analyse

"In total 947 participants were recruited with a median age of 22 years (interquartile range (IQR) 8-51) and a male/female ratio of 0.84. 289 out of 947 (30.5%) had loiasis according to our case definition and positivity for RAPLOA was the most frequent loiasis-defining characteristic (65.1%; 188/289)
(Table 2). Prevalence of pruritus was 32.3% (306/947) in the overall study population and 67.2% (636/947) reported being regularly engaged in activities that involved exposure to the forest at least twice
weekly. The sub-study assessing Calabar swelling recruited 213 participants (median age 16; IQR 7 - 40, male/female ratio of 0.90). 60/213 (28.2%) had loiasis."

```{r echo = FALSE, warning = FALSE, message = FALSE}

lois <- read.dta13("data/dbloias.dta")

n_partic <- 
  lois  %>%
  mutate(np = str_detect(id,"id")) %>%
  count(np)  %>%
  mutate(variable = "Participant",
         res = as.character(n)) %>% 
  select(variable, res)


med_age <- 
  lois %>% 
  summarise(med = median(age),
            q1 = quantile(age, 0.25),
            q2 = quantile(age, 0.75),
            res = paste0(med, " (", q1, " - ", q2, ")")) %>% 
  mutate(variable = "Median age (range)") %>% 
  select(variable, res)


loisdef <- 
  lois  %>%
  count(eyewormmf) %>% 
  mutate(perc = round( n * 100 / sum(n), 1),
         res = paste0(n, " (", perc, "%)"),
         variable = case_when(eyewormmf == "No" ~ "Loiasis (No)",
                              eyewormmf == "Yes" ~ "Loiasis (Yes)")) %>% 
  select(variable, res) %>%
  slice(2)

pos_raplo <- 
  lois  %>%
  count(mf,eyeworm) %>% 
  slice(2:4) %>% 
  mutate(perc = round( n * 100 / sum(n), 1),
         res = paste0(n, " (", perc, "%)"),
         variable = case_when(eyeworm == "Yes" ~ "RAPLOA +")) %>% 
  select(variable, res) %>%
  slice(1)

pruri <- 
  lois  %>%
  group_by(itch) %>%
  summarise(nb = n()) %>%
  mutate(perc = round(nb*100 / sum(nb),1),
         res = paste0(nb, " (", perc, "%)"),
         variable = "Prevalence of pruritus") %>%
  select(variable, res)  %>%
  slice(2)

twice <- 
  lois  %>%
  group_by(anyexposure) %>%
  summarise(nb = n()) %>%
  mutate(perc = round(nb*100 / sum(nb),1),
         res = paste0(nb, " (", perc, "%)"),
         variable = "Exposure to forest") %>%
  select(variable, res) %>%
  slice(2)

calabsw_recr <- 
  lois %>% 
  mutate(calabsw = str_detect(edema, "edema")) %>% 
  count(calabsw) %>%
  mutate(variable = "Calabar swelling recruted",
         res = as.character(n)) %>% 
  select(variable, res) %>%
  slice(1)

calablois <- 
  lois %>%
  filter(edema %in% c("Yes","No")) %>%
  count(eyewormmf)  %>%
  mutate(perc = round(n * 100 / sum(n), 1),
          res = paste0(n, " (", perc, "%)"),
        variable = " Calabar swelling with loiasis") %>%
  select(variable, res) %>%
  slice(2)

partext <- bind_rows(n_partic, med_age, loisdef, pos_raplo, pruri,
                     twice,calabsw_recr, calablois) %>%
  rename(Variables = variable , CEDARO = res)


kable(partext, caption = "Section 1 Results")

```


```{r echo = FALSE, message = FALSE, warning = FALSE}



options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)


lois <-
  lois %>%
  mutate(age_group = case_when(
    age < 6 ~ '< 6',
    age >= 6 & age <= 17  ~ '6-17',
    age >= 18  & age <= 34  ~ '18-34',
    age >= 35 & age <= 49  ~ '35-49',
    age >= 50 & age <= 64  ~ '50-64',
    age >= 65 ~ '65+',
    TRUE ~ 'other'))
med_age <- 
  lois %>% 
  summarise(med = median(age),
            q1 = quantile(age, 0.25),
            q2 = quantile(age, 0.75),
            res = paste0(med, " (", q1, " - ", q2, ")")) %>% 
  mutate(variable = "Median age (range)") %>% 
  select(variable, res)


med_age2 <- 
  lois %>% 
  filter(eyewormmf == "Yes") %>%
  group_by(eyewormmf) %>%
  summarise(med1 = median(age),
            q3 = quantile(age, 0.25),
            q4 = quantile(age, 0.75),
            res = paste0(med1, " (", q3, " - ", q4, ")")) %>% 
  mutate(variable = "Median age (range)") %>%
  select(variable, res)

med_age3 <- 
  lois %>% 
  filter(edema %in% c("No","Yes")) %>% 
  summarise(med = median(age),
            q3 = quantile(age, 0.25),
            q4 = quantile(age, 0.75),
            res = paste0(med, " (", q3, " - ", q4, ")")) %>% 
  mutate(variable = "Median age (range)") %>% 
  select(variable, res)

age <- bind_cols(med_age, med_age2, med_age3) %>%
  select(variable,res,res1,res2) %>%
  rename(v1 = variable, v2 = res, v3 = res1, v4 = res2)

sex1 <- 
  lois  %>%
  count(sex) %>% 
  mutate(perc = round( n * 100 / sum(n), 1),
         sex_loi = paste0(n, " (", perc, "%)"))  %>%  
  select(sex, sex_loi) %>%
  slice(2)

sex2 <-  
  lois %>%
  filter(eyewormmf == "Yes") %>%
  count(sex) %>%
  mutate(perc = round(n * 100 / sum(n), 1),
         sex_loi = paste0(n, " (", perc, "%)")) %>%
  select(sex, sex_loi) %>%
  slice(2)

sex3 <-  
  lois %>%
  filter(edema %in% c("Yes","No")) %>%
  count(sex) %>%
  mutate(perc = round(n * 100 / sum(n), 1),
         sex_loi = paste0(n, " (", perc, "%)")) %>%
  select(sex, sex_loi) %>%
  slice(2)

sex <- bind_cols(sex1, sex2, sex3)  %>%
    select(sex, sex_loi, sex_loi1, sex_loi2) %>%
    rename(v1 = sex, v2 = sex_loi, v3 = sex_loi1, v4 = sex_loi2)

cohoage <- 
  lois  %>%
  count(age_group) %>%
  mutate(perc = round(n *100 / sum(n), 1),
         res3 = paste0(n, " (", perc, "%)"),
         var3 = age_group) %>%
  select(var3, res3)


loisage <- 
  lois  %>%
  filter(eyewormmf == "Yes") %>%
  count(age_group) %>%
  mutate(perc = round(n *100 / sum(n), 1),
         res2 = paste0(n, " (", perc, "%)"),
         var2 = age_group) %>%
  select(var2, res2)

cabage <- 
  lois  %>%
  filter(edema %in% c("Yes","No")) %>%
  count(age_group) %>%
  mutate(perc = round(n *100 / sum(n), 1),
         res1 = paste0(n, " (", perc, "%)"),
         var1 = age_group) %>%
  select(var1, res1)

p1 <- bind_cols(cohoage, loisage, cabage) %>%
     select(var3, res3, res2, res1) %>%
  rename(v1 = var3, v2 = res3, v3 = res2, v4 = res1)

p2 <- bind_rows(age, p1)

expo1 <-  
  lois %>%
  count(exposureforest) %>%
  mutate(perc = round(n * 100 / sum(n),1),
         cohort = paste0(n, " (", perc, "%)")) %>%
  select(exposureforest, cohort)

expo2 <-  
  lois %>%
  filter(eyewormmf == "Yes") %>%
  count(exposureforest, eyewormmf) %>%
  mutate(perc = round(n * 100 / sum(n),1),
         loiasis = paste0(n, " (", perc, "%)")) %>%
  select(exposureforest, loiasis)

expo3 <-  
  lois %>%
  filter(edema %in% c("Yes","No")) %>%
  count(exposureforest) %>%
  mutate(perc = round(n * 100 / sum(n),1),
         calaba = paste0(n, " (", perc, "%)")) %>%
  select(exposureforest, calaba)

 p3 <- bind_cols(expo1, expo2, expo3) %>%
   select(exposureforest, cohort,loiasis,calaba) %>%
   rename(v1 = exposureforest, v2 = cohort, v3 = loiasis, v4 = calaba)


raplo1 <- 
  lois %>%
  count(eyeworm) %>%
  mutate(perc = round(n * 100 / sum(n), 1),
         res1 = paste0(n, " (", perc, "%)")) %>%
  select(eyeworm, res1) %>%
  slice(2)

raplo2 <- 
  lois %>%
  filter(eyeworm == "Yes") %>%
  count(eyewormmf) %>%
  mutate(perc = round(n * 100 / sum(n), 1),
         res2 = paste0(n, " (", perc, "%)"),
         variable = "Positive") %>%
  select(variable, res2)


raplo3 <- 
  lois %>%
  filter(edema %in% c("Yes", "No")) %>%
  count(eyeworm) %>%
  mutate(perc = round(n * 100 / sum(n), 1),
         res3 = paste0(n, " (", perc, "%)"),
         variable = "Positive") %>%
  select(variable, res3) %>%
  slice(2)

raplo <- bind_cols(raplo1, raplo2, raplo3) %>%
   select(variable, res1, res2, res3) %>%
  rename(v1 = variable, v2 = res1, v3 = res2, v4 = res3)

mf1 <- 
  lois  %>%
  count(mf) %>%
  mutate(perc = round(n * 100 / sum(n), 1),
         res4 = paste0(n, " (", perc, "%)"),
         variable = "Positive") %>%
  select(variable, res4) %>%
  slice(2)

mf2 <- 
  lois %>%
  filter(eyewormmf == "Yes") %>%
  count(mf) %>%
  slice(2) %>%
  mutate(perc = round(n * 100 / sum(n), 1),
         res5 = paste0(n, " (", perc, "%)")) %>%
  select(mf, res5)

mf3 <- 
  lois %>%
  filter(edema %in% c("Yes", "No")) %>%
  count(mf) %>%
  mutate(perc = round(n * 100 / sum(n), 1),
         res6 = paste0(n, " (", perc, "%)")) %>%
  select(mf, res6) %>%
  slice(2)

microfilarea <- bind_cols(mf1, mf2, mf3) %>%
  mutate(v1 = "Positive") %>%
  rename(v2 = res4, v3 = res5, v4 = res6) %>%
 select(v1, v2, v3, v4) 

rapmf1 <- 
  lois  %>%
  count(eyewormmf) %>%
  slice(2) %>%
  mutate(perc = round(n * 100 / sum(n), 1),
         res7 = paste0(n, " (", perc, "%)"),
         variable1 = eyewormmf)  %>%  
  select(variable1, res7)

raplomicrofilarea <- bind_cols(loisdef,rapmf1, calablois) %>%
       mutate(v1 = "positive") %>%
       rename(v2 = res, v3 = res7, v4 = res1) %>%
       select(v1, v2, v3, v4) 

prupos2 <- 
  lois  %>%
  filter(eyewormmf == "Yes") %>%
  count(itch) %>%
  mutate(perc = round(n * 100 / sum(n), 1),
         res8 = paste0(n, " (", perc, "%)")) %>%
  select(itch, res8) %>%
  slice(2)

prupos3 <- 
  lois %>%
  filter(edema %in% c("Yes", "No")) %>%
  count(itch) %>%
  mutate(perc = round(n * 100 / sum(n), 1),
         res9 = paste0(n, " (", perc, "%)")) %>%
  select(itch, res9) %>%
  slice(2)

pruritus <- bind_cols(pruri,prupos2, prupos3) %>%
  mutate(v1 = "positive") %>%
  rename(v2 = res, v3 = res8, v4 = res9) %>%
 select(v1, v2, v3, v4) 

tab1 <- bind_rows(p2,p3, sex, raplo)
tab2 <- bind_rows(microfilarea, raplomicrofilarea,pruritus)
tab3 <- bind_rows(tab1, tab2) %>%
  rename("Characteristics" = v1, "No.(column %)" = v2, "No.(row %)" = v3,
         "No.(column %)" = v4)

```

\begin{landscape}

```{r echo = FALSE, warning = FALSE}


tab3 <- tab3[c(1, 2, 6, 3:5, 7, 12, 8:11, 13:16),]

rownames(tab3) <- NULL


tab3$Characteristics <- c("Median (IQR)","below 6" ,"6-17","18-34","35-49","50-64","65+","Male","None","Occasional",
  "Frequent","Intensive","Positive","Positive","Positive","Positive")

kable(tab3, "latex",longtable=T, booktabs = T, caption ="Section 1 results", align = "c") %>%
kable_styling("striped", full_width = F) %>%
add_header_above(c(" ","Total cohort (N = 947)","Loiasis (N = 289)",
                   "Population of Calabar swelling sub-study (N = 213)")) %>%
pack_rows("Age(years)", 1, 7, label_row_css = "background-color: #666; color: #fff;") %>%
pack_rows("Sex",8,8,label_row_css = "background-color: #666; color: #fff;") %>%
pack_rows("Exposure to forest",9, 12, label_row_css = "background-color: #666; color: #fff;") %>%
  pack_rows("RAPLO", 13,13, label_row_css = "background-color: #666; color: #fff;") %>%
pack_rows("Microfilareamia", 14,14, label_row_css = "background-color: #666; color: #fff;") %>%
pack_rows("RAPLO or microfilareamia", 15,15, label_row_css = "background-color: #666; color: #fff;") %>%
pack_rows("Pruritus", 16,16, label_row_css = "background-color: #666; color: #fff;")  
 

```
\end{landscape}



- CEDARO ID: NA
- Date of report: 2019-10-21
- Analysis performed by: Eddy MBENA
- Analysis supervised by: Fabrice Mougeni & Bertrand Lell
- Raw data supplied by: Frieder Schaumburg
- Code repository: https://gitlab.com/MBENA/loiasis
- Local storage path: NAS

